---
title: "rRNADepletion_StrippedSerum"
author: "Lachlan Baer"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    fig_width: 8
    fig_height: 6
    fig_align: "center"
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  fig.align = "center"
)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# Setup 

```{r packages}
library(tidyverse)
library(magrittr)
library(parallel)
library(here)
library(scales)
library(AnnotationHub)
library(Biostrings)
library(Rsamtools)
```

```{r options}
if (interactive()) setwd(here::here())
theme_set(theme_bw())
cores <- detectCores() - 2
```

```{r loadObjects}
load(here::here("4_T47D_ZR75_DHT_StrippedSerum/R/output/4_1_DE.RData"))
load(here::here("4_T47D_ZR75_DHT_StrippedSerum/R/output/4_2_kmer.RData"))
```

```{r ah}
ah <- AnnotationHub() %>%
  subset(species == "Homo sapiens") %>%
  subset(rdataclass == "EnsDb")
ensDb <- ah[["AH83216"]]
```

# Exploration

Now that differential expression of $k$-mers has been detected, the next step is to develop a normalisation method that can lessen the extent of gene-level differential expression bias due to presence of particular $k$-mers. The initial aim of this exploration is to define a single value per gene, $\kappa$, that represents the magnitude of bias due to the genes constituent $k$-mers.

## Using logFC

In an initial attempt to develop a value for $\kappa$, logFC values from previously derived $k$-mer differential expression testing will be evaluated. Firstly, for each gene's coding sequence, a count for the presence of all detected $k$-mers will be calculated. This count, $n$, will then be multiplied by the $k$-mers logFC value to determine $\kappa$: \[\kappa = n_1LFC_1 + n_2LFC_2 + n_3LFC_3\]

```{r ids}
ids <- res_r$Geneid
```

```{r}
## Coding sequences were saved to speed up document
cds <- exonsBy(ensDb, by = "gene", filter = GeneIdFilter(ids)) %>%
  lapply(function(x){
    GenomicRanges::reduce(x)
  }) %>%
  GRangesList()
saveRDS(cds, here::here("4_T47D_ZR75_DHT_StrippedSerum/R/output/cds.Rds"))
cds <- readRDS(here::here("4_T47D_ZR75_DHT_StrippedSerum/R/output/cds.Rds"))
```

```{r}
seqs <- mclapply(
  cds, 
  function(x){
    getSeq(
      FaFile(
        file = "/hpcfs/users/a1647910/refs/ensembl-release-101/homo_sapiens/Homo_sapiens.GRCh38.dna.primary_assembly.fa", 
        index = "/hpcfs/users/a1647910/refs/ensembl-release-101/homo_sapiens/Homo_sapiens.GRCh38.dna.primary_assembly.fa.fai"), 
      x
    ) %>%
      unlist() 
  },
  mc.cores = cores
)
```

### k = 5

```{r}
k5mers <- k5topTable$mer
```

```{r}
# k5list <- mclapply(
#   seqs,
#   function(x){
#     vapply(
#       k5mers,
#       function(y){countPattern(y, x)},
#       numeric(1)
#     ) 
#   },
#   mc.cores = cores
# )
# saveRDS(k5list, here::here("4_T47D_ZR75_DHT_StrippedSerum/R/output/k5list.Rds"))
k5list <- readRDS(here::here("4_T47D_ZR75_DHT_StrippedSerum/R/output/k5list.Rds"))
```

```{r}
k5bias <- mclapply(
  k5list,
  function(x){
    x %>%
      enframe(name = "mer", value = "n") %>%
      left_join(k5topTable) %>%
      dplyr::select(mer, n, logFC) %>%
      mutate(bias = n * logFC) %>%
      .$bias %>%
      sum()
  },
  mc.cores = cores
) %>%
  unlist() %>%
  enframe(name = "Geneid", value = "bias")
## 2176.81 sec elapsed
```

```{r}
k5bias %>%
  left_join(res_r) %>%
  mutate(bias = bias / maxLen) %>%
  ggplot(aes(bias, -log10(P.Value))) +
  geom_point()

k5bias %>%
  left_join(res_r) %>%
  ggplot(aes(bias, log10(maxLen))) +
  geom_point() +
  geom_smooth(method = "lm")
```

### k = 7

```{r}
k7mers <- k7topTable$mer
```

```{r}
k7list <- mclapply(
  seqs,
  function(x){
    vapply(
      k7mers,
      function(y){countPattern(y, x)},
      numeric(1)
    )
  },
  mc.cores = cores
)
saveRDS(k7list, here::here("4_T47D_ZR75_DHT_StrippedSerum/R/output/k7list.Rds"))
k7list <- readRDS(here::here("4_T47D_ZR75_DHT_StrippedSerum/R/output/k7list.Rds"))
k7bias <- mclapply(
  k7list,
  function(x){
    x %>%
      enframe(name = "mer", value = "n") %>%
      left_join(k7topTable) %>%
      dplyr::select(mer, n, logFC) %>%
      mutate(bias = n * logFC) %>%
      .$bias %>%
      sum()
  },
  mc.cores = cores
) %>%
  unlist() %>%
  enframe(name = "Geneid", value = "bias")
```

```{r}
k7bias %>%
  left_join(res_r) %>%
  mutate(bias = bias / maxLen) %>%
  ggplot(aes(log10(maxLen), -log10(P.Value))) +
  geom_point()
```

```{r}
k7bias %>%
  dplyr::rename(k7bias = bias) %>%
  left_join(k5bias) %>%
  ggplot(aes(k7bias, bias)) +
  geom_point()
```